\Huge Conferences in the Era of Expensive Carbon

### 2020-02-10 main.tex
Notes

\usepackage{xcolor}
\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}
\newcommand{\bcp}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf bcp: #1]}}\fi}
\newcommand{\mwh}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf mwh: #1]}}\fi}
\newcommand{\COtwoE}{CO$_2$e}

@@

A broad scientific consensus warns that human
emissions of greenhouse gases are warming the earth.
% (Figure \ref{fig:warming}).
This is a present-day emergency: the UN's Intergovernmental
Panel on Climate Change (IPCC) says a 40\% decrease in
emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC18}.
% (Figure~\ref{fig:IPCC-cover}).
Reductions on this scale
require urgent and sustained commitment at all levels of society---not only
national, state, and city governments, but also
universities, companies, and scientific societies.

@@

Indeed, scientific societies have an especially important role to
play, since, for many members, travel to conferences represents a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.8 tons of carbon dioxide (\COtwoE, or informally
``carbon'') per
passenger~\cite{carboncalculator}. This is a significant fraction of the
{\em total} yearly emissions for an average resident of the US (16.5 tons)
or Europe (7 tons)~\cite{emissions}.
Moreover, these emissions have no near-term technological
fix, as jet fuel is difficult to replace with renewable energy
sources~\cite{elec-air}.

How should ACM respond to these facts?

In 2016, ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to consider this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport}, we are putting
forward two concrete proposals. First, all ACM
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on---in particular, from travel to
the conference. Second, ACM should put a \emph{price} on
carbon in conference budgets, creating a steady pressure on organizers to reduce
their footprints.

@@

Our first proposal is modest: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint}.
%
These reports should be collected in a central place, in a uniform format.
Most conferences' footprints will be dominated by participants' air travel,
but the data gathered should go beyond this to include ground
transportation, travel to in-person program committee meetings, and
estimated emissions from hotels and food.

ACM should develop tools to gather and publicize this data.  For example,
SIGPLAN recently built an air-travel-focused carbon calculator for
conferences~\cite{ClimateCommittee}.  Users can upload conference
registration data, and the calculator will estimate the {\COtwoE} cost of
air travel for those sufficiently far away.
% This calculator could be
% expanded to account for other forms of conference emissions.

There is some the reputational risk to ACM in taking the step of publicizing
its carbon footprint: the numbers are likely to be high, and they may be
used to criticize both the ACM and the broader academic community.  But
making this information available is a crucial first step: we cannot manage
what we have not measured.

\PARAGRAPH{Easy trimming}
One effect of public accounting will be to nudge conference organizers and
attendees to change their behavior. By analogy, chain restaurants in the US
are now required by law to post calorie counts of food items on menu boards;
studies show that enlightened customers order, on average, up to 50 fewer
calories a day~\cite{menu}.

Similarly, SIGPLAN has been considering how to reduce emissions, informed by
an accounting of its own carbon footprint. This discussion has led the
organizers of two flagship
conferences (POPL and ICFP) to switch from in-person to online program
committee meetings, joining a trend among other SIGs, and has prompted
several conferences to increase investments in livestreaming and video
recording to support remote participation.

\begin{figure}[t!]
\centering
\includegraphics[width=6in]{SIGPLAN-confs2.png}
\includegraphics[width=6in]{ParticipantsOriginAll2.png}
% \includegraphics[width=6in]{Distances.png}
\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
  conferences.  The smallest dot (ICFP
  14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
  largest (ICFP 16, in Nara, Japan) represents 1.94 tons per
  participant. Bottom: Breakdown of continent-of-origin for participants in
  each conference.  Colored bars represent percentages of participants whose
  home city is in each continent: blue for North America, orange for Europe,
  green for Asia.
% \bcp{Richard: I find this a
%     little hard to follow. Maybe you could simply label the different groups
%     of conferences by their location? I think the "(NA)", "(EU)", and
%     "(IAS)" are a bit hard to read and decode at first, so that I didn't
%     understand where each conference was located at first glance. (Adrianne
%     had the same response when I showed her.) Maybe just put text under
%     each, maybe with a horizontal bracket, that says "North America,"
%     "Europe," and "Rest of world," with the bracket and/or words color-coded
%     as well.}\fi
%
% Bottom: Distances traveled by participants, one way, in Km; middle dashed
% line is the median travel distance.
}
\label{fig:confs}
\end{figure}

@@

However, while public accounting of emissions will encourage easy
reductions, it is not likely, by itself, to induce major shifts in behavior.
%
Science is a fundamentally social process, and the conference system
accelerates scientific research through high-bandwidth interaction,
direct dissemination of results, network building, and serendipitous
cross-fertilization.
%
Organizers and attendees will naturally be reluctant to consider changes
that might threaten these benefits.

@@

To illustrate the challenges, consider the problem of choosing a conference
location that minimizes emissions from participants' travel.  Using recent
registration data from four \mbox{SIGPLAN} conferences, Figure~\ref{fig:confs}
shows two ways of looking at the relation between locations and emissions.
%
The top diagram shows an
estimated per-participant {\COtwoE} footprint for each instance of each
conference over the past ten years (excluding a few for which we
had trouble getting data), with larger dots representing higher emissions.
Eyeballing this diagram, it might seem that carbon-conscious
organizers should hold all of these conferences in either the northeast US or
western Europe every year.  But the bottom visualization tells a
different story.  The horizontal colored bar at the bottom represents
the continent on which each conference was held, and each vertical bar gives
a breakdown of the participants in that conference, colored according to the
continent of their work address.  A glance
at the colors makes clear that, though a minority come from far away, the
majority of participants in each conference are local to the region where the
conference is being held.
%
This suggests that always locating conferences in the
same one or two places would significantly impact the diversity of the
research community by discouraging participation from other parts of the
world.  Indeed, one might conclude that, from the point of view of
strengthening the research community, conferences should move around as
much as possible!
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate perspectives suggest that significantly reducing conference
emissions may
require genuinely painful compromises.  The impulse to ignore the issue is
entirely understandable.

@@

\SECTION{Put a Price on Carbon}
%
This dilemma is a microcosm  of one faced by all of society. To address it,
many policy experts advocate using some form of carbon pricing to impose a
concrete, immediate cost on
emissions~\cite{carbonprice}. Doing so makes manifest the hidden
environmental cost of emissions, incentivizing
{\COtwoE}-reducing changes without mandating exactly which ones, and thus
allowing for creative and efficient responses. Continuing the junk-food
analogy,
some municipalities including Berkeley and Philadelphia have imposed a
per-calorie tax on soft drinks; studies found that doing so
significantly reduced consumption~\cite{sodatax}.

Thus, our second proposal is that {\bf ACM should impose a surcharge on
  conferences based on their carbon footprint.}  The charge should start low
and increase steadily and predictably, year on year.
%
Conference organizers can then choose how best to balance their
budgets---whether
by decreasing per-participant emissions, decreasing (physical)
participation, increasing registration fees, soliciting corporate
sponsorship, or other means.
In this process, a primary concern
should be to find ways of {reducing} the financial burden of such a
surcharge on those disproportionately affected by it---students
without grant support, participants from developing areas, etc.
Well-funded participants should subsidize the carbon surcharges of less
wealthy ones.

Ideally, at some point, governments will impose carbon pricing uniformly,
and all carbon-intensive activities will have to pay it. But ACM can send a
strong message about the importance of this issue---and get ahead of the
coming changes---by acting now.

@@

\PARAGRAPH{Precipitating change}
%
What should ACM do with the funds collected from this surcharge?  One
obvious possibility is purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor, who uses the funds to finance an
activity that permanently removes or avoids emitting some amount of
greenhouse gases.  The veracity and permanence of this activity is certified
by a watchdog organization. (E.g., planting trees is often considered not to
be a certifiable
activity, since it's hard to guarantee they won't be cut down;
reductions from
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities are
easier to predict.)
Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
academic societies, and even energy companies such as Exxon, now use
carbon offsets to reduce their
net footprint. ACM conferences should consider doing the same, and the
purchases should be
included in the public accounting we are proposing.
% For example\bcp{beter
%   transition?}, ACM recently
% declared that conferences can offer attendees the option to pay for an
% offset during registration.  Other scientific societies and conferences are
% going further, rolling a carbon offset into the registration fee for every
% participant.

Beyond buying offsets, one can
imagine many good uses for the funds generated from a carbon surcharge: defraying the costs of
virtualizing conferences (livestreaming, etc.), and supporting ``green''
computing research~\cite{rolnick2019tackling}.  As an example
of the last, ACM could help fund a
cross-cutting research initiative specifically aimed at understanding how to
best replace or approximate the socializing and networking aspects of
conferences in a virtual setting.

Ultimately, however, carbon offsets and other ``good works'' cannot
substitute for real reductions in emissions~\cite{anderson2012inconvenient,
  carbon-offsets-are-not-our-get-out-jail-free-card}: they are, at best, a
short-term expedient that buys time to agree on more
difficult cuts.
%
Indeed, the main goal of carbon pricing should be to stimulate creative
rethinking of the conference model itself---for example, seriously
considering alternatives such as rapid-turnaround journal-only publishing
models, yearly megaconferences, and entirely virtual
conferences~\cite{NCN}.
%
A potential sticking point is that some of these will significantly reduce
conference revenues, in turn impacting the income stream of ACM itself; this
could make emissions reduction politically problematic unless ACM's
conference-focused business model is also adjusted.

@@

Another key issue in implementing this second proposal will be how to
set a price that reflects the true social cost of carbon, without unduly
harming the scientific community.
%
Initial data gathering will play a role, and a steady and predictable annual
increase is a necessary component, but setting both the initial
price and the slope of the ramp will likely be difficult political
decisions.
%
One measurable target, in line with the latest evidence from climate
science, would be to tune the parameters with a goal of halving actual
emissions every decade from now on, following a recently proposed ``Carbon
Law''~\cite{Rockstrom1269}.

@@

The climate crisis is too urgent to leave to world leaders to address at
their own pace: Organizations at every scale, including ACM, must confront
their own contributions, raise awareness and foster discussion among their
members~\cite{acm-climate}, and establish new ways of doing business in the
lower-carbon future that is now upon us.  We in ACM should do our part by
mandating public accounting of conference carbon footprints and by putting a
concrete price on the carbon that we use.

@@

@@

\iflater

\iflater
\bcp{Things still missing:
  \begin{itemize}
  \item do we need any more citations for the claims we make?
  \item Find a place for the {\tt acm-climate} mailing list~\cite{acm-climate}
  \item Recheck and fix citations
  \item scan the stuff below and delete it
  % \item Write a cover letter
  % \begin{itemize}
  % \item We tried hard to remove incrementalism and make it much more
  % hard hitting
  % \item In particular, we make two concrete proposals
  % \item However, we did keep the suggestion of carbon offsets, hopefully
  % in a more convincing / palatable / better argued form
  % \item what else?
  % \end{itemize}
  \end{itemize}
}
\fi

\newpage
\section*{Permissions for figures}

@@

If you've already thought about all of these, then that's great.  If not, please do take them as a set of potential
challenges to consider (all voluntary... as I intend them only as suggestions).
\end{quote}
\item
About Crista's figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item
Also about Crista's figure:
\begin{itemize}
\item
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I'm looking…
\item
Consider citing the new IEA report as another piece of evidence that action is needed
\item
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item
Is there a ``more authoritative'' source for price of carbon offsets than
our carbon report?
\item
Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\end{itemize}
\fi

>>>>

\newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
\newcommand{\mwh}[1]{\ifdraft{\bf [mwh: #1]}\fi}
\newcommand{\COtwoE}{CO$_2$e}

@@

A broad scientific consensus warns that human
emissions of greenhouse gases are warming the earth.
% (Figure \ref{fig:warming}).
This is not a problem that can be left to future
generations: The UN's Intergovernmental
Panel on Climate Change (IPCC) says a 40\% decrease in
emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC18}.
% (Figure~\ref{fig:IPCC-cover}).
Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society---not only
national, state, and city governments, but also 
universities, companies, and scientific societies.

@@

Indeed, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger\iflater\bcp{Citation?}\fi, a significant fraction of the
{\em total} yearly emissions for an average resident of the US (16.5 tons)
or Europe (7 tons)~\cite{emissions}.  
Moreover, these emissions have no near-term technological
fix, since jet fuel is difficult to replace with renewable energy
sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  Two years ago, \mwh{Three, now?} ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to research this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport} and intensive and
wide-ranging discussions, we are putting forward two concrete
proposals. First, all ACM 
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on---in particular, from travel to
the conference. Second, ACM should put a \emph{price} on 
carbon in conference budgets, to create pressure on organizers to reduce
their footprints.

@@

Our first proposal is modest: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint}.  
%
Public accounting will nudge conference organizers and attendees to change
their behavior. By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.  
%MWH: I don't know why this should be a parenthetical.  Why would the
%analogy be an aside? It is meant to connect the reader's understanding to
%what we are proposing, explciitly; it is meant to be read, not skipped
%over. 
% BCP: In general, parenthetical remarks are not meant to be skipped over:
% they are meant to be read as asides.  I think this reads better in parens,
% but I won't fight for it.

As a case in point, SIGPLAN has been considering how to reduce emissions
over the past couple of years, informed by an accounting of its own
{\COtwoE} expenditures. This discussion has led the organizers of two of its
flagship conferences (POPL and ICFP) to switch from in-person to on-line
program committee meetings (joining a trend among other SIGs). This
discussion has also prompted several conferences to increase investments in
livestreaming and video recording to support remote participation.

\PARAGRAPH{Carbon offsets}
%
SIGPLAN has also explored how to mitigate conference emissions by purchasing
\emph{carbon offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor who uses the funds to finance an
activity that permanently removes or avoids emitting some amount of
greenhouse gases.  The veracity and permanence of this activity is certified
by a watchdog organization. (Typically, planting trees is not a certifiable
activity, since it's hard to guarantee the trees won't be cut down, but
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities can
be.)

Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
academic societies, and even energy companies such as Exxon, now use
carbon offsets in order to reduce their 
net footprint. ACM conferences can do the same, and the purchases can be
mentioned in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration.  Other scientific societies are going further,
rolling carbon offsets into the registration fees for all
participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi

We acknowledge that carbon offsets are a 
  controversial 
  tool~\cite{anderson2012inconvenient,
  carbon-offsets-are-not-our-get-out-jail-free-card}.  Offsets are a
useful short-term expedient, not a replacement for more difficult
cutting. 
% MWH: It's a bit odd to have two lengthy parenthetical asides in the same
% paragraph. I think this one is on point, so need not be an aside.
% BCP: The whole paragraph was in parens, before.  However, the new paragraph
% header does a decent job of "closing the aside," so I agree the parens can
% go.

% \mwh{Should
% we mention the FCRC numbers here? 45 out of 2,700 used {\COtwoE} offsets}
% That's getting too much down into the details, I think -- also, these
% numbers reflect a hastily organized, preliminary version of what we really
% want to do.

\PARAGRAPH{Data gathering}
%
Beyond encouraging conferences to trim their emissions, public 
accounting will generate useful data for future organizers.  
The data should be made
available in 
a central place, in a uniform format.  Most
conferences' footprints will be dominated by participants' air travel, but the
data gathered should go beyond this to include air travel to in-person PC
meetings and estimated hotel and food {\COtwoE} footprints.

ACM should develop tools to gather and publicize this data. 
SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
 the right one? Link seems
 dead!  But Richard says ``As a reader, my first question is, how do I get
 access to this calculator? Can you cite this calculator in some way? It
 would be great to give a citation to where people can learn about it or
 use it, even if just, "To learn more about the calculator, e-mail xxx"''}\fi. 
Users can upload conference registration data, and the
calculator will estimate the {\COtwoE} cost of air travel for those
sufficiently far away. This calculator could be expanded to account for
other forms of conference emissions.

We acknowledge that there is some the reputational risk to ACM in taking the
step of publicizing its carbon footprint: the numbers are likely to be high,
and may be used to criticize both the ACM and the broader academic
community.  Nevertheless, making this information available is a crucial
first step: we cannot manage what we have not measured.

\SECTION{Put a Price on Carbon}

While public accounting of {\COtwoE} expenditures will incentivize some 
reductions, it is not likely, by itself, to induce major changes.
%
Science is a fundamentally social process, and the conference system
accelerates scientific research through high-bandwidth interaction,
direct dissemination of results, network building, and serendipitous
cross-fertilization.
%
Organizers and attendees will be reticent to make changes that might
threaten these benefits.

\begin{figure}[t]
\centering
\includegraphics[width=6in]{SIGPLAN-confs2.png}
\includegraphics[width=6in]{ParticipantsOriginAll2.png}
% \includegraphics[width=6in]{Distances.png}
\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
  conferences.  The smallest dot (ICFP
  14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
  largest dot (ICFP 16, in Nara, Japan) represents 1.94 tons per
  participant. Bottom: Breakdown of continent-of-origin for participants in
  each conference.  Colored bars represent percentages of participants whose
  home city is in each continent.  \iflater\bcp{Richard: I find this a
    little hard to follow. Maybe you could simply label the different groups
    of conferences by their location? I think the "(NA)", "(EU)", and
    "(IAS)" are a bit hard to read and decode at first, so that I didn't
    understand where each conference was located at first glance. (Adrianne
    had the same response when I showed her.) Maybe just put text under
    each, maybe with a horizontal bracket, that says "North America,"
    "Europe," and "Rest of world," with the bracket and/or words color-coded
    as well.}\fi 
% Bottom: Distances traveled by participants, one way, in Km; middle dashed
% line is the median travel distance. 
}
\label{fig:confs}
\end{figure}

@@

@@

To illustrate the challenges, consider the problem of choosing a conference
location that minimizes emissions from participants' travel.  Using recent
registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
shows two ways of looking at the relation between locations and emissions. 
%
The top diagram shows an
estimated per-participant {\COtwoE} footprint for each instance of each
conference over the past ten years (excluding a few for which we
had trouble getting data), with larger dots representing more emissions.
Eyeballing this diagram, it seems that carbon-conscious 
organizers should hold all of these conferences in either the northeast US or
western Europe every year.  But the bottom visualization tells a
different story.  The horizontal colored bar at the bottom represents
the continent on which each conference was held, and each vertical bar gives
a breakdown of the participants in that conference, colored according to the
continent of their work address.  From a superficial glance
at the colors, it is clear that, though a minority come from far away, the
majority of participants in each conference are local to the region where the
conference is being held.\iflater\bcp{That point could be strengthened by adding some
  numbers (just in the text, not a figure) about what percentage of
  conference attendees attend even one additional instance of the same
  conference...}\fi{} This suggests that locating
conferences every year in the same one or two places would significantly
damage the 
research community by discouraging participation from other parts of the
world.\bcp{Richard: The damage to the research community is a little unclear to me on first read. I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world". Or perhaps some better world, but more specificity of the damage.}  Indeed, one might conclude that, from the point of view of
strengthening the research community, conferences should move around as 
much as possible.
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}  
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate perspectives suggest that reducing conference emissions will
require genuinely painful compromises.  The impulse to avoid thinking about
it is entirely understandable!

@@

\PARAGRAPH{A cost for carbon}
%
This dilemma is a microcosm of one faced by all of society. To address it,
many policy experts advocate some form of carbon pricing (either taxes or
cap-and-trade schemes), imposing a concrete, immediate cost on
emissions~\cite{carbonprice}. Doing so makes manifest the hidden
environmental cost of emissions, incentivizing
{\COtwoE}-reducing changes without specifying exactly which, and thus
allowing for creative and efficient responses. Continuing the junk-food
analogy, some 
municipalities such as New York city and Philadelphia have imposed a
per-calorie tax on soft drinks, with some studies finding that doing so
indeed reduced consumption~\cite{sodatax}.
%MWH: The above is not a parenthetical, so the earlier junkfood analogy should not be, either

Thus, our second proposal is that {\bf ACM should impose a surcharge on
  conferences based on their carbon footprint.}  The charge should start low 
and increase steadily and predictably, year on year.
%
Conference organizers can then choose how best to balance their
budgets---whether 
by decreasing per-participant emissions, decreasing (physical)
participation, increasing registration fees, soliciting additional corporate
sponsorship, or other means.

Ideally, at some point, governments will impose carbon pricing uniformly,
and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
strong message about the importance of this issue---and get ahead of the
coming changes---by acting now.

@@

\PARAGRAPH{Precipitating change}
 %
What should ACM do with the funds collected from this surcharge?  One can
imagine many good uses, including carbon offsetting, defraying the costs of
virtualizing conferences (livestreaming, etc.), and supporting ``green''
computing research. As an example of the last, ACM could help fund a 
cross-cutting research initiative specifically aimed at understanding how to
best replace or approximate the socializing and networking aspects of
conferences in a virtual setting.  But the very first concern
should be to find ways of {\em reducing} the financial burden of such a
surcharge on those who may be disproportionately affected by it---students
without grant support, participants from developing areas, etc.  That is,
wealthier participants should subsidize the carbon footprints of less
wealthy ones.

Carbon pricing will stimulate creative rethinking of the conference model 
itself---for example, seriously considering alternatives such as
rapid-turnaround, journal-only publishing models, once-a-year
megaconferences, and entirely virtual conferences~\cite{NCN}.
%
A potential sticking point is that some of these will significantly reduce
conference revenues, in turn impacting the income stream of ACM itself; this
could make emissions reduction politically problematic unless ACM's
conference-focused business model is also adjusted.

@@

A key issue in implementation of this proposal will be how to set a price
that reflects the true social cost of carbon, without unduly harming the
community.
%
Initial data gathering will play a role, and a steady and predictable annual
ramp on the price is a necessary component, but setting both the initial
price and the slope of the ramp will likely be difficult political
decisions.
%
One measurable target, in line with the latest evidence from climate
science, would be to halve emissions every decade from now on, following the
``Carbon Law'' proposed in a recent paper~\cite{Rockstrom1269}.  \bcp{Read the
  paper and cite it!
  https://doi.org/10.1126/science.aah3443
  }

@@

The climate crisis is too urgent to leave for world leaders to address at
their own pace: Organizations at every scale, including ACM, must confront
their own contributions, raise awareness among their members, and establish
new ways of doing business in the lower-carbon future that is coming at us.

@@

\iflater
\bcp{Things still missing:
  \begin{itemize}
  \item do we need any more citations for the claims we make?
  \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
  \end{itemize}
}
\fi

@@

\iflater
\newpage
\section*{Permissions for figures}

@@

If you've already thought about all of these, then that's great.  If not, please do take them as a set of potential
challenges to consider (all voluntary... as I intend them only as suggestions).
\end{quote}
\item 
About Crista's figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item 
Also about Crista's figure:
\begin{itemize}
\item 
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item 
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I'm looking…
\item 
Consider citing the new IEA report as another piece of evidence that action is needed
\item 
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item 
Is there a ``more authoritative'' source for price of carbon offsets than
our carbon report?
\item 
Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\end{itemize}
\fi
